### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Done for tonight, I moved a lot of things around for Chapter 6, and I wrote a few paragraphs detailing more things. I need to finish by Thursday evening, and maybe Friday launch new experiments to finish appendix.

% Regret minimization in stochastic bandits has been extensively studied since the works of \cite{Robbins52} and \cite{LaiRobbins85}, and several algorithms with a $\cO(\log(T))$ \emph{problem-dependent} regret upper bound have been proposed (see, \eg, \cite{LattimoreBanditAlgorithmsBook} for a survey). Among those, the \klUCB{} algorithm (\cite{KLUCBJournal}) has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
% \emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.

@@

\paragraph{Outline.}
%
This chapter is structured as follows.
We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}. In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on the regret of both variants.
The unified regret analysis for these two algorithms is given in Section~\ref{sec:6:RegretAnalysis}.
Numerical experiments are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.

@@

Introducing the number of change-points on arm $i$, defined as $\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.

@@

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To the authors knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.

@@

\paragraph{Advantages of our approach.}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
This study is supported by both theoretical and empirical results.
%
Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing  much in terms of regret.

@@

Sequential change-point detection has been extensively studied in the statistical community (see, \eg, \cite{Basseville93} for a survey). In this section, we are interested in detecting changes on the mean of a probability distribution with bounded support. Assuming that  we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$. We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a stopping time $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.

@@

The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, one can show that $\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right]$,
where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
%
This motivates the following definition of the Bernoulli GLRT change point detector.

@@

% ----------------------------------------------------------------------------
\subsection{Non-Asymptotics Properties of the B-GLRT}\label{subsec:6:PropGLR}

When used for a bandit problem, the two main properties of a sequential change-point detection test are its false alarm probability and its detection delay.
%
A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arms will be detected after a short enough amount of samples from that arm.
%
We give below two Lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.

@@

In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small. To define $\beta$, we need to introduce the function $\cT$, defined for $x>0$ by
%
\begin{equation}\label{def:6:function_T}
    \cT(x) ~:=~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
\end{equation}
where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$.
%
And for any $x \ge 0$, $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise. The function $\cT$ is easy to compute numerically.
The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96} (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
%
Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given in Appendix~\ref{proof:6:FalseAlarm}.

@@

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{equation}
    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
\end{equation}
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{}-like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
the number of selections of arm $i$, and
\[\widehat{\mu}_i(t) = (1/n_i(t)) \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ (if $t>1$ else $f(t)=0$),
the index is defined as
%
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
\end{equation}

\paragraph{Two options.}
%
In this work, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
%
Under local restart, in the general case the times $\tau_i(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
%
One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq. $(31)$-$(32)$.}:
$f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.

@@

The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by (\cite{CaoZhenKvetonXie18}), whereas \CUSUMUCB{} performs randomized exploration.
%
We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
%
A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples}, proved below.

\begin{proposition}\label{prop:6:EnoughSamples}
    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that $n_i(t) - n_i(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor$.
\end{proposition}

@@

%-----------------------------------------------------------------------------
\subsection{A Unified Regret Analysis}

\TODOL{Rewrite, include the proof of at least one case here! No need to be that quick for the thesis, we only wrote it like this for the COLT version \cite{Besson2019GLRT}.}

@@

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$. Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
\begin{align*}
 &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}
%
Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
%
\begin{align*}
    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{align*}
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
%
To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

@@

>>>>

% Regret minimization in stochastic bandits has been extensively studied since the works of \cite{Robbins52} and \cite{LaiRobbins85}, and several algorithms with a $\cO(\log(T))$ \emph{problem-dependent} regret upper bound have been proposed (see, \eg, \cite{LattimoreBanditAlgorithmsBook} for a survey). Among those, the \klUCB{} algorithm \cite{KLUCBJournal} has been shown to be asymptotically optimal for Bernoulli distributions (in that it exactly matches the lower bound given by \cite{LaiRobbins85}) and can also be employed when the rewards are assumed to be bounded in $[0,1]$.
% \emph{Problem-independent} upper bounds of the form $R_T=\Omega(\sqrt{KT})$ (with no hidden constant depending on the arms distributions) have also been established for stochastic algorithms, like MOSS, or \klUCB-Switch by \cite{Garivier18}, while \klUCB{} is known to enjoy a sub-optimal $\cO(\sqrt{KT\ln(T)})$ problem-independent regret.

@@

\paragraph{Outline.}
%
% This chapter is structured as follows.
We introduce the model and review related works in Section~\ref{sec:6:BanditSetting}. In Section~\ref{sec:6:ChangePointDetector}, we study the Generalized Likelihood Ratio test for sub-Bernoulli distributions (B-GLRT) as a Change-Point Detector (CPD) algorithm.
We introduce the two variants of \GLRklUCB{} algorithm in Section~\ref{sec:6:GLRklUCB_Algorithm}, where we also present upper bounds on the regret of both variants.
The unified regret analysis for these two algorithms is given in Section~\ref{sec:6:RegretAnalysis}.
Numerical experiments are presented in Section~\ref{sec:6:NumericalExperiments}, with more details in the Appendix.

@@

Introducing the number of change-points on arm $i$, defined as $\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.

@@

\paragraph{Illustrations of two problems of our benchmark.}\label{par:6:benchmark1}

To illustrate the setup we presented above, we give here two problems of our benchmark, used for the numerical experiments presented below in Section~\ref{sec:6:NumericalExperiments}.

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.

@@

\paragraph{Advantages of our approach.}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
This study is supported by both theoretical and empirical results presented in this chapter.
%
In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{i,t} |\mu_i(t) - \mu_i(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{i,j} |\mu_i(t) - \mu_j(t)|$.

Moreover, we can use the two problems $1$ and $2$ to illustrate the expected empirical behavior of our proposal \GLRklUCB.
Problem $1$ has only \emph{local changes}, meaning that at any change-point, only one arm sees its distribution change, while problem $2$ has only \emph{global changes}, meaning that all arms see their distribution change at each change-point.
They illustrate both the extreme cases of having $C_T = \Upsilon_T$ (for problem $1$) and $C_T = K \Upsilon_T$ (for problem $2$).
\begin{itemize}
    \item
On one hand, we expect the \emph{local restart} variant of \GLRklUCB{} to outperform the \emph{global restart} variant for problem $1$, as the intuition suggests that it is sub-optimal to reinitialize the memory of the observations of the $K-1$ arms whose distributions did not change after a change-point is detected (as the \emph{global restart} variant do).
    \item
On the other hand, we expect the \emph{global restart} variant to outperform the \emph{local restart} variant for problem $2$, as detecting a change one any arm is enough to know that all arms changed and to reinitialize the memory of all arms, and thus the intuition suggests that it is sub-optimal to reinitialize only the memory of the observation of the arm on which the change-point was detected (as the \emph{local restart} variant do).
\end{itemize}
%
Of course, on a given problem, without additional prior knowledge on the difficulty of the problem at hand, the algorithm cannot know which situation is more likely to happen, only local changes ($C_T = \Upsilon_T$) or only global changes ($C_T = K \Upsilon_T$) or any intermediate setting between the two extreme cases.
Thus we do not believe to be able to design a policy that could be uniformly better than the two variants of our proposal \GLRklUCB.
%
Surprisingly, our numerical experiments show that the \emph{local} variant of \GLRklUCB{} actually outperforms the \emph{glocal} variant for both problem $1$ and $2$, as shown below in Table~\ref{table:6:totalResults1}, and in other problems as shown in Table~\ref{table:6:totalResults2}.
%
Understanding this difference in terms of numerical performance of the two variants is left as future work.

Finally, on the practical side, while we can note that the proposed GLR test is more complex to implement than the test used by \MUCB, we propose two heuristics to speed it up while not losing much in terms of regret, in Appendix~\ref{sub:6:IdeasOptimizations}.

@@

Sequential change-point detection has been extensively studied in the statistical community.
We refer to the book \cite{Basseville93} for a survey.
%
In this section, we are interested in detecting changes on the mean of a probability distribution with bounded support.

Assume that we collect independent samples $X_1,X_2,\ldots$ all from some distribution supported in $[0,1]$.
We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\ldots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\ldots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a stopping time $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.

@@

The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl}, we can show that in any one-dimensional exponential family, we have
\[\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right],\]
%
where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
%
This motivates the following definition of the Bernoulli GLRT change point detector.

@@

% ----------------------------------------------------------------------------
\subsection{Non-Asymptotic Properties of the B-GLRT}\label{subsec:6:PropGLR}

When used for a bandit problem, the two main properties of a sequential change-point detection test are its false alarm probability and its detection delay.
%
A small false alarm probability ensures that no detection occurs before it should (\ie, no useless detection occur on stationary segments), and a small detection delay ensures that, if the stationary segments are long enough, then every change-points on every arms will be detected after a short enough amount of samples from that arm.
%
We give below two lemmas, Lemma~\ref{lem:6:FalseAlarm} which bounds the false alarm probability for a choice of threshold function, and Lemma~\ref{lem:6:Delay} which bounds the detection delay if there is enough samples before a change.

@@

In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small. To define $\beta$, we need to introduce the function $\cT$, defined for $x>0$ by
%
\begin{equation}\label{def:6:function_T}
    \cT(x) ~:=~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
\end{equation}
where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$.
%
And for any $x \ge 0$, $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise. The function $\cT$ is easy to compute numerically.
The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96} (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
%
Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
% The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given below.

@@

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{equation}
    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
\end{equation}
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{}-like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
the number of selections of arm $i$, and
\[\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ (if $t>1$ else $f(t)=0$),
the index is defined as
%
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
\end{equation}

\paragraph{Two options.}
%
For this algorithm, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
%
Under local restart, in the general case the times $\tau_i(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
%
One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq. $(31)$-$(32)$.}:
$f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.

@@

The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by \cite{CaoZhenKvetonXie18}, whereas \CUSUMUCB{} performs randomized exploration.
%
We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
%
A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples}, proved below.

\begin{proposition}\label{prop:6:EnoughSamples}
    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $i$ (\ie, for a $k\in\{1,\dots,\NCi\}$, one has $\tau_i^{(k)} = \tau_i(t) < s \leq t < \tau_i^{(k+1)}$) it holds that
    \begin{equation}
        n_i(t) - n_i(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor.
    \end{equation}
\end{proposition}

@@

%-----------------------------------------------------------------------------
\subsection{Sketch of the unified regret analysis}

@@

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$. Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
\begin{align*}
 &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}
%
Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
%
\begin{align}\label{eq:6:GeneRegretBound}
    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{align}
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
%
To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

@@

\TODOL{Justify MORE AND WAY MORE about all these results!}
